Learning to Synthesize Rules for Network Intrusion Detectors

### 2020-01-15 paper/paper.tex
pass on 4.3.3

Similarly to the the previous heuristic functions, this function uses
the frequency of content modifiers to calibrate the contribution of
each rule. More precisely, we compute this function with the following
formula:

\[\sum_{i=1}^{N}\frac{\sum_{j=1}^{M_i}\frac{\mathit{modifier\_frequency[modifier_{ij}}]}{\mathit{max(modifier\_frequency)}}}{M_i}\]

\noindent
, where $N$ is the number of contents in a given rule and $M_i$ is the
number of content modifiers that could access the string associated
with content $i$. \Mar{Qual o rationale para considerar as varias
  alternativas de acesso ao inves de apenas a alternativa que eh mencionada?}%% Isto
                                                                             %% nao
                                                                             %% intuitivo
                                                                             %% para
                                                                             %% mim
                                                                             %% -Marcelo

@@

\begin{figure}[t!]
  \centering
  \scalebox{1.2}{
    \begin{tikzpicture}
      \begin{axis}[
          ybar stacked,
          enlargelimits=0.15,
          legend style={at={(0.5,-0.15)},
            anchor=north,legend columns=-1},
          %          ylabel={\#rules},
          symbolic x coords={http\_method, http\_uri, http\_user\_agent, http\_header, http\_host, http\_connection},
          xtick=data,
          xticklabel style={rotate=45},          
          nodes near coords,
          every node near coord/.append style={font=\scriptsize},        
          nodes near coords align={vertical},
        ]
        \addplot coordinates {(http\_method,3821) (http\_uri,10111) (http\_user\_agent,1022) (http\_header,4089) (http\_host,224) (http\_connection,29)};
      \end{axis}
    \end{tikzpicture}
  }
  \vspace{-2ex}
  \caption{\label{fig:distribution-content_modifiers}Content modifiers
    frequency.\Mar{what about ``none'' (no modifier)?}}
\end{figure}

Consider a rule containing the option \CodeIn{content:
  ``/HtmlAdaptor''; http\_uri}, as the one from
Section~\ref{sec:jboss}. The contribution of this option to the score
of the corresponding rule would be $0.5=\frac{10,111}{10,111*1}$. The
value of $M_i$ for this case is 1 as the field of the HTTP message
\CodeIn{Request URI} holding the string can only be accessed through
the modifier \CodeIn{http\_uri}. For the option \CodeIn{content:
  ``keep-alive''; <mod>}, either \CodeIn{http\_header} or
\CodeIn{http\_connection} could substitute the placeholder
\CodeIn{<mod>} to locate the string ``keep-alive'' stored in the field
\CodeIn{Connection} in the header of an HTTP request
message\Mar{Lucas, a imagem mostrando header com campos Request URI e
  Connection e os conteudos citados acima pode ajudar.}. Consequently,
for this case, the constribution of content modifiers would be
$0.2=\frac{4089+29}{10111*2}$. Assuming that both options are
mentioned in a rule, the aggregated contribution of this heuristic
function would be \Fix{...}\Mar{Lucas, confirme que nao esta faltando
  calcular a media (dividir tudo por N).}

>>>>

Similarly to the the previous heuristic functions, this function uses
the frequency of content modifiers to calibrate the score of each
rule. More precisely, we compute this function with the following
formula:

\[\frac{\sum_{i=1}^{N}g(i)}{N},~g(i)=\frac{\sum_{j=1}^{M_i}\frac{\mathit{modifier\_frequency[modifier_{ij}}]}{\mathit{max(modifier\_frequency)}}}{M_i}\]

\noindent
, where $N$ is the number of contents options, $g(i)$ is the
contribution of content option $i$ to the score,
$\mathit{modifier\_frequency[modifier_{ij}}]$ denotes the popularity
of a modifier $j$ to content option $i$,
$\mathit{max(modifier\_frequency)}$ refers to the modifier with
maximum popularity (for normalization), and $M_i$ is the number of
content modifiers that could access the string associated with content
$i$.

%% \Mar{Qual o rationale para considerar as varias
%%   alternativas de acesso ao inves de apenas a alternativa que eh mencionada?}0
%%   \Luc{Vc está supondo que as regras que geramos incluem os modificadores e que deveriamos pontuar somente de acordo com os que estão presentes? Não eh o caso. A ideia aqui basicamente eh: se um certo content do payload está numa posição q eh muito verificada nas regras, possivelmente esse content eh mais relevante do q outro q está numa posição q raramente eh verificada. Para avaliar isso nós só precisamos saber quantas vezes os modificadores que acessam essas posições aparecem no dataset.}
%%   \Luc{Uma coisa que pensei agora foi de considerar apenas o modificador mais comum entre aqueles que acessam a posição de um content. Como a pontuação de um content eh dado pela media das ocorrencias dos modificadores que conseguem acessá-lo, se houver um modificador muito comum e varios modificadores raros a pontuação dada será baixa.}

@@

\begin{figure}[t!]
  \centering
  \scalebox{1.2}{
    \begin{tikzpicture}
      \begin{axis}[
          ybar stacked,
          enlargelimits=0.15,
          legend style={at={(0.5,-0.15)},
            anchor=north,legend columns=-1},
          %          ylabel={\#rules},
          symbolic x coords={http\_method, http\_uri, http\_user\_agent, http\_header, http\_host, http\_connection, http\_request\_line},
          xtick=data,
          xticklabel style={rotate=45},          
          nodes near coords,
          every node near coord/.append style={font=\scriptsize},        
          nodes near coords align={vertical},
        ]
        \addplot coordinates {(http\_method,3821) (http\_uri,10111) (http\_user\_agent,1022) (http\_header,4089) (http\_host,224) (http\_connection,29) (http\_request\_line,11)};
      \end{axis}
    \end{tikzpicture}
  }
  \vspace{-2ex}
  \caption{\label{fig:distribution-content_modifiers}Content modifiers
    frequency.}
\end{figure}

%\Mar{what about ``none'' (no modifier)?}}
%  \Luc{Uma regra pode sim ter contents sem modificadores, mas isso n
    %% vem ao caso aqui. Nessa heuristica toda string do payload vai ter
    %% associada a ela os modificadores q podem acessar a sua
    %% posição. Portanto, não vai existir o caso de não haver
    %% modificador. Atenção: isso não quer dizer q as regras q geramos
    %% incluem os modificadores, elas n incluem.

Consider a rule containing the option \CodeIn{content:
  ``/HtmlAdaptor''}. The contribution of this option to the score of
the corresponding rule would be $0.5$
(=$\frac{10,111}{10,111*1}$). The value of $M_i$ for this case is 1 as
the field of the HTTP message \CodeIn{Request URI}---holding the
string ``/HtmlAdaptor''---can be accessed through the modifier
\CodeIn{http\_uri}. Conceptually, the relatively high heuristic value
given to this option indicates that this is a reasonable option
because the string used in the content option (``/HtmlAdaptor'') is
stored in a specific field of the HTTP message (\CodeIn{Request URI})
and can be accessible through content modifiers
(\CodeIn{http\_uri}). For the option \CodeIn{content: ``keep-alive''},
either \CodeIn{http\_header} or \CodeIn{http\_coxnnection} could be
used as modifier to locate the string ``keep-alive'' stored in the
field \CodeIn{Connection} (see Figure~\Fix{xxx}). Consequently, for
this case, the contribution of modifiers to this content option would
be $0.2$ (=$\frac{4089+29}{10111*2}$). Assuming that both options are
used in a rule, the aggregated contribution of this heuristic function
would be 0.35 (=(0.2+0.5)/2).
